---
title: 'Topic 4: Plotting of Likelihhod function and MLE'
author: "Shubhangi Mistari"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



**Q.1**
Generate a random sample of size 50 from $Poisson(\lambda=2.8)$ distribution. Plot a likelihood function against different values of $\lambda$
and check whether likelihood function attains maxima at 
$1)\bar{X}$
2) 2.8
```{r}
set.seed(123)
n=50;lambda=2.8 
x=rpois(n,lambda)
lambda_=seq(1.5,4.5,length.out=500) #different values of lambda
like_func= function(lambda_) {
  exp(-n*lambda_)*lambda_^sum(x)
}
plot(lambda_,like_func(lambda_),main="Likelihood function plot for poisson distribution",
     xlab = "values of lambda_",ylab="Likelihood function")
abline(v=mean(x),col="black")
abline(v=2.8,col="green")
cat("MLE of lambda is",mean(x),"\n")


neg_like_func= function(lambda_) { 
  -exp(-n*lambda_)*lambda_^sum(x)
}

nlm(neg_like_func,2.5)

optimize(f=neg_like_func,interval = c(0,5))

optimize(f=like_func,interval = c(0,5),maximum = T)


```

Log likelihood 
```{r}
set.seed(123)
n=50;lambda=2.8 
x=rpois(n,lambda)
lambda_=seq(1.5,4.5,length.out=500) #different values of lambda
log_like_func= function(lambda_) {
  -n*lambda_+sum(x)*log(lambda_)
}
plot(lambda_,log_like_func(lambda_),main="Log likelihood function plot for poisson distribution",
     xlab = "values of lambda_",ylab="Log likelihood function")
abline(v=mean(x),col="black")
abline(v=2.8,col="green")
cat("MLE of lambda is",mean(x),"\n")


neg_log_like_func= function(lambda_) {
   -(-n*lambda_+sum(x)*log(lambda_))
}
nlm(neg_log_like_func,2.5)

optimize(f=neg_log_like_func,interval = c(0,5))

optimize(f=log_like_func,interval = c(0,5),maximum = T)

```


**Q.2**
Generate a random sample of size 50 from $Burnoulli(p=0.6)$ distribution. Plot a likelihood function against different values of $p$
and check whether likelihood function attains maxima at 
$1)\bar{X}$
2) 0.6
```{r}
set.seed(123)
n=50;p=0.6 
x=rpois(n,p)
p_=seq(0.3,0.9,length.out=500) #different values of lambda
like_func= function(p_) {
  p_^sum(x)*(1-p_)^(n-sum(x))
}
plot(p_,like_func(p_),main="Likelihood function plot for Burnoulli distribution",
     xlab = "values of p_",ylab="Likelihood function")
abline(v=mean(x),col="black")
abline(v=0.6,col="green")
cat("MLE of p is",mean(x),"\n")

```




**Q.3**
Generate a random sample of size 50 from $N(\mu=4.6,\sigma^2=1)$ distribution. Plot a likelihood function for
1) Different values of$\mu$ when $\sigma=1$ known and check whether likelihood function attains maxima at 
a) $\bar{X}$
b)  4.6
```{r}
set.seed(123)
mu=4.6
sigma2=1
n=50
x=rnorm(50,4.6,1)
mu_=seq(3,6,length.out=500)
likelihood= function(mu_){
  exp(-0.5*((n*mu_^2)-(2*mu_*sum(x))))
}
plot(mu_,likelihood(mu_),main="Likelihood function plot for Normal dist (sigma known)",
     xlab = "values of mu_",ylab=" Likelihood function")
abline(v=mean(x),col="black")
abline(v=4.6,col="green")
cat("MLE of mu is",mean(x),"\n")
```


2) Different values of$\sigma$ when $\mu=4.6$ known and check whether likelihood function attains maxima at
a)$1/n\sum(Xi-\bar{X})$
b)$1$
```{r}
set.seed(123)
mu=4.6;sigma2=1;n=50
x=rnorm(50,4.6,1)
sigma_2=seq(0,01.5,length.out=1000)
likelihood= function(sigma_){
  ((1/(2*pi*sigma_2))^n/2)*exp(-(0.5/sigma_2)*sum((x-4.6)^2))
}
plot(sigma_2,likelihood(sigma_2),main="Likelihood function plot for Normal dist (mu known)",
     xlab = "values of mu_",ylab=" Likelihood function")
abline(v=((n-1)/n)*var(x),col="pink")
abline(v=(1/n)*sum((x-4.6)^2),col="black")
abline(v=1,col="green")
cat("MLE of sigma2 is (population variance) is :",1/n*sum((x-4.6)^2),"\n")
```

**Q.4**
Generate a random sample of size 50 from $Exponential(mean=\theta=15.2)$ distribution. Plot a likelihood function against different values of $\theta$ and check whether likelihood function attains maxima at 
$1)\bar{X}$
2) 15.2
```{r}
set.seed(123)
theta=15.2
n=50
x=rexp(50,1/15.2)
theta_=seq(5,30,length.out=500)
likelihood= function(theta_){
  (1/theta_)^n*exp(-sum(x)/theta_)
}
plot(theta_,likelihood(theta_),main="Likelihood function plot for Exp(mean=15.2) dist ",
     xlab = "values of theta_",ylab=" Likelihood function")
abline(v=mean(x),col="black")
abline(v=15.2,col="green")
cat("MLE of theta is",mean(x),"\n")
```





**Q.5**
Consider the following random sample of size 10 Generated from $U(0,\theta), \theta>0$ distribution. 
2.06874,5.82498,2.65698,3.58650,2.78196,1.36005,4.49083,5.74037,3.41361,1.36519.  Plot a likelihood function and cheak
whether likelihood function attains maxima at 
a) $2\bar{X}$
b) $X_{(n)}$
```{r}
theta=seq(3,8,length.out=1000)
x=c(2.06874,5.82498,2.65698,3.58650,2.78196,1.36005,4.49083,5.74037,3.41361,1.36519)
likelihood=function(theta){
  (1/theta)^n*ifelse(max(x)<theta,1,0)
}
plot(theta,likelihood(theta),main="Likehood function plot for uniform distribution",
     xlab="theta",ylab="Likelihood function")
abline(v=2*mean(x),col="red")
abline(v=max(x),col="black") 
cat("mle of theta is",max(x),"\n")
```






**Q.6**
Generate a random sample of size 50 from $Cauchy(\mu,1)$ distribution. Use Newton Raphson method and method of scoring to find MLE of $\mu$.

```{r}
#newton raphson 
set.seed(12)
n=50;mu=2.8
x=rcauchy(n,mu,1)

v=c(median(x),rep(0,9)) #9 for 10 iterations


for(i in 1:length(v)){
  y=x-v[i]
  d_11=2*sum(y/(1+y^2))     #dlogL/dmu
  d_12=-2*sum(1-y^2/(1+y^2)^2)    
  v[i+1]=v[i]-d_11/d_12
  
}
print(v)
```


```{r}
#Method of scoring
set.seed(12)
n=50;mu=2.8
x=rcauchy(n,mu,1)

u=c(median(x),rep(0,9)) #9 for 10 iterations

I_x=1/2

for(i in 1:length(u)){
  y=x-u[i]
  d_1=2*sum(y/(1+y^2))    #dlogL/dmu
  u[i+1]=u[i]+d_1/n*I_x
  
}
u
```


**Q.7**
Generate a random sample of size 50 from $E(\theta,1)$ distribution. Plot a likelihood function against different values of $\theta$ and check whether likelihood function attains maxima at 
a) $\bar{X}-1$
b) $X_{(1)}$
```{r}
set.seed(123)
theta=1.1
n=50
theta_=seq(0.8,1.2,length.out=1000)
y=runif(n,0,1)
x=theta-log(1-y)

likelihood= function(theta_){
  exp(-sum(x)+n*theta_)*ifelse(theta_<min(x),1,0)
}
plot(theta_,likelihood(theta_),main="Likehood function plot for Exp(theta,1) distribution",
     xlab="theta",ylab="Likelihood function")
abline(v=mean(x)-1,col="red")
abline(v=min(x),col="black")
cat("mle of theta is",min(x),"\n")
```

**Q.8**
Generate a random sample of size 50 from $Normal(\mu,1)$ distribution. Use Newton Raphson method and method of scoring to find MLE of $\mu$.

```{r}
# Newton Raphson
set.seed(123)
n=50;mu=1
x=rnorm(n,mu,2)
v=c(median(x),rep(0,9)) #9 for 10 iterations

for(i in 1:length(v)){
  
  d_11=n*(mean(x)-v[i] )    #dlogL/dmu
  d_12=-n
  v[i+1]=v[i]-d_11/d_12
  
}
print(v)
```

```{r}
#Method of scoring
set.seed(12)
n=50;mu=2.8
x=rnorm(n,mu,1)

u=c(median(x),rep(0,9)) #9 for 10 iterations

I_x=1

for(i in 1:length(u)){
  
  d_1=n*(mean(x)-u[i]) #dlogL/dmu
  u[i+1]=u[i]+d_1/n*I_x
  
}
u
```




















